---
title: "Table Formatting"
format: html
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.show = "hold", 
                      time_it = TRUE, dpi = 100)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = T, collapse = TRUE)

library(terra)
library(lme4)
library(MASS)
library(lmerTest)
library(MuMIn)
library(ggplot2)
library(ggeffects)
library(ggtext)
library(merTools)
library(glmnet)
library(stats)
library(ggcorrplot)
library(forcats)
library(RColorBrewer)
library(webshot)
library(kableExtra)
library(tidyverse)
library(flextable)
library(ggnewscale)
library(tidyterra)
library(ggpubr)
library(ggspatial)
library(patchwork)
library(randomForest)
library(mlr3)
library(mlr3pipelines)
library(pdp)
library(kernelshap)
library(shapviz)
library(here)
here()
#knitr::knit_hooks$set(webgl = hook_webgl)
rgl::setupKnitr(autoprint = TRUE)

setGDALconfig("GDAL_PAM_ENABLED", "FALSE")

#tmap_options(raster.max.cells = c(plot = 1e7, view = 1e6))
```

Data sources
```{r} 
#| include: false
#| echo: false
#| message: false
#| warning: false
#| cache: true
# knitr::purl("/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/analysis/WA_SOC_controls/PNW_SOC_Prediction.qmd", output = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/analysis/WA_SOC_controls/PNW_SOC_Prediction.R")

source(here("analysis/All_WA_Functions.R"))
 
```

### Models

```{r}

lmm_stocks <- readRDS(here("analysis/models/PNW_Model_Dredge_Spline.rds"))
lmm_stocks_noWIP <- readRDS(here("analysis/models/PNW_Model_Dredge_Spline_noWIP.rds"))
lmm_lab <- readRDS(here("analysis/models/PNW_Model_Lab_dredge_best.rds"))

QRFM_stocks <- readRDS(here("analysis/models/MLR3_RFM_Ranger_tuned.rds"))
QRFM_stocks_noWIP <- readRDS(here("analysis/models/MLR3_RFM_Ranger_tuned_noWIP.rds"))
QRFM_lab <- readRDS(here("analysis/models/MLR3_RFM_Ranger_tuned_LabData.rds"))

```

### Dataframes

```{r}
#| cache: true


All_WIP_Acc_DF <- read.csv(here("data/dataframes/All_WIP_Acc_DF.csv"))
wipPredictors <- read.csv(here("data/dataframes/PNW_WIP_Predictors.csv"))
allsite_WIPimp <- read.csv(here("data/dataframes/PNW_WIP_VarImp.csv"))

pnw_dat <- read.csv(here("data/dataframes/PNW_data_MLR3_RFM_LMM.csv"))
wa_stocks_dat <- read.csv(here("data/dataframes/All_WA_1m_stocks_preds.csv"))

train <- read.csv(here("data/dataframes/Train_data.csv")) |> 
  dplyr::select(-X) |> dplyr::mutate(across(where(is.character), as.factor))
test <- read.csv(here("data/dataframes/Test_data.csv")) |> 
  dplyr::select(-X) |> dplyr::mutate(across(where(is.character), as.factor)) |> 
  mutate(geomorphons = factor(geomorphons, levels = levels(train$geomorphons)))
test_viz <- read.csv(here("data/dataframes/Test_data.csv")) |> 
  dplyr::select( -X) |> dplyr::mutate(across(where(is.character), as.factor)) |> 
  mutate(geomorphons = factor(geomorphons, levels = levels(train$geomorphons)))
test.wet <- read.csv(here("data/dataframes/RFM_Test_Wet_data.csv")) |> 
  dplyr::select(-X) |> dplyr::mutate(across(where(is.character), as.factor)) |> 
  mutate(geomorphons = factor(geomorphons, levels = levels(train$geomorphons)))
test.upl <- read.csv(here("data/dataframes/RFM_Test_Upl_data.csv")) |> 
  dplyr::select(-X) |> dplyr::mutate(across(where(is.character), as.factor)) |> 
  mutate(geomorphons = factor(geomorphons, levels = levels(train$geomorphons)))
RF_testpred <- read.csv(here("data/dataframes/PNW_SOC_RF_test_pred_response.csv")) |> 
  select(-WIP) |> 
  cbind(test)
LMM_testpred <- read.csv(here("data/dataframes/PNW_SOC_LMM_test_pred_response.csv")) |> 
  cbind(test)
RF_testprednw <- read.csv(here("data/dataframes/PNW_SOC_RF_noWIP_test_pred_response.csv")) |> 
  cbind(test)
LMM_testprednw <- read.csv(here("data/dataframes/PNW_SOC_LMM_noWIP_test_pred_response.csv")) |> 
  cbind(test)

pnw_dat_lab <- read.csv(here("data/dataframes/PNW_Lab_data_MLR3_RFM_LMM.csv")) |> 
  dplyr::select( -X)
trainlab<- read.csv(
          file = here("data/dataframes/Train_Lab_data.csv")) |> 
  dplyr::select(-X)
testlab<- read.csv(
          file = here("data/dataframes/Test_Lab_data.csv"))|> 
  dplyr::select(-X)
testlab_viz<- read.csv(
          file = here("data/dataframes/Test_Lab_data.csv"))|> 
  dplyr::select( -X)
testlab_wet<- read.csv(
          file = here("data/dataframes/RFM_Test_Lab_Wet_data.csv"))|> 
  dplyr::select(X)
testlab_upl<- read.csv(
          file = here("data/dataframes/RFM_Test_Lab_Upl_data.csv")) |> 
  dplyr::select(-X)


train_dat_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Train_Scale.csv"))
train_dat_wet_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Train_Wet_Scale.csv"))
train_dat_upl_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Train_Upl_Scale.csv"))
test_dat_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Test_Scale.csv"))
test_dat_wet_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Test_Wet_Scale.csv"))
test_dat_upl_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Test_Upl_Scale.csv"))

trainlab_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Train_Scale.csv"))
trainlab_wet_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Train_Wet_Scale.csv"))
trainlab_upl_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Train_Upl_Scale.csv"))
testlab_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Test_Scale.csv"))
testlab_wet_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Test_Wet_Scale.csv"))
testlab_upl_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Test_Upl_Scale.csv"))

anova_ci_stocks <- read.csv(paste0(here("data/dataframes/anova_best_dredge_model_confint95.csv")))
anova_ci_stocks_nw <- read.csv(paste0(here("data/dataframes/anova_best_dredge_model_confint95_noWIP.csv")))
anova_ci_lab <- read.csv(paste0(here("data/dataframes/LMM_Lab_confint95.csv")))


PI_RMSE_stocks <- readRDS(paste0(here("data/dataframes/LMM_PredictInterval_RMSE_Stocks.rds")))
PIw_RMSE_stocks <- readRDS(paste0(here("data/dataframes/LMM_Wet_PredictInterval_RMSE_Stocks.rds")))
PIu_RMSE_stocks <- readRDS(paste0(here("data/dataframes/LMM_PredictInterval_Upl_RMSE_Stocks.rds")))

PI_RMSE_lab <- readRDS(paste0(here("data/dataframes/LMM_PredictInterval_RMSE_Lab.rds")))
PIw_RMSE_lab <- readRDS(paste0(here("data/dataframes/LMM_Wet_PredictInterval_RMSE_Lab.rds")))
PIu_RMSE_lab <- readRDS(paste0(here("data/dataframes/LMM_PredictInterval_Upl_RMSE_Lab.rds")))

mod_errdf <- read.csv(here("data/dataframes/PNW_Stocks_model_fit.csv"))
mod_errdf_noWIP <- read.csv(here("data/dataframes/PNW_Stocks_model_fit_noWIP.csv"))
mod_errdf_lab <- read.csv(here("data/dataframes/PNW_Lab_model_fit.csv"))
qrf_errdf <- read.csv(here("data/dataframes/MLR3_RFM_ErrorDF.csv"))
qrf_errdf_noWIP <- read.csv(here("data/dataframes/MLR3_RFM_ErrorDF_noWIP.csv"))
qrf_errdf_lab <- read.csv(here("data/dataframes/MLR3_RFM_ErrorDF_LabData.csv"))


ks <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap.rds")) 
ks_wet <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_Wet.rds")) 
ks_upl <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_Upl.rds")) 
ks_nw <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_noWIP.rds"))

ia_1 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Interact_1_plot.rds")) 
ia_1$data <- ia_1$data |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_2 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_2_plot.rds")) 
ia_2$results <- ia_2$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_3 <- readRDS( 
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_3_plot.rds")) 
ia_3$results <- ia_3$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_4 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_4_plot.rds")) 
ia_4$results <- ia_4$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_5 <- readRDS( 
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_5_plot.rds"))  
ia_5$results <- ia_5$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)

ks_lab <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_LabData.rds")) 
ks_lab_wet <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_Wet_LabData.rds")) 
ks_lab_upl <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_Upl_LabData.rds")) 

ia_lab_1 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Interact_Lab_1_plot.rds")) 
ia_lab_1$data <- ia_lab_1$data |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_lab_2 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_Lab_2_plot.rds")) 
ia_lab_2$results <- ia_lab_2$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_lab_3 <- readRDS( 
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_Lab_3_plot.rds")) 
ia_lab_3$results <- ia_lab_3$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_lab_4 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_Lab_4_plot.rds")) 
ia_lab_4$results <- ia_lab_4$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_lab_5 <- readRDS( 
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_Lab_5_plot.rds")) 
ia_lab_5$results <- ia_lab_5$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)

```
Map tables

```{r}
#| cache: true


PNW_SOC_df <- read.csv(here("data/dataframes/PNW_MapSOC_wetupl_1m_no_wip_no_wip.csv")) |> 
  mutate(across(where(is.numeric) & !dplyr::contains("Total_area"), \(x) x * 100))

nwca_soc_est <- read.csv(here("data/dataframes/NWCA_SOC_C_map_simp.csv")) |> 
  mutate(StudyArea = str_extract(Name, "Hoh|Mas|Col"),
         Name = str_remove_all(Name, "Hoh |Mas |Col ")) |> 
  dplyr::select(StudyArea, everything()) |>
  mutate(StudyArea = case_when(str_detect(StudyArea, "(?i)HOH") ~ "HRW",
                            str_detect(StudyArea, "(?i)MAS") ~ "MRW",
                            str_detect(StudyArea, "(?i)COL") ~ "CVW",
                            str_detect(StudyArea, "(?i)HLEF") ~ "HLEF",
                            .default = NA))

mcnicol_soc_frac_form <- read.csv(here("data/dataframes/McNicol_2019_SOC_Frac_Estimates_HLEF.csv"))

gsoc_frac_pnw <- read.csv(here("data/dataframes/GSOC_Frac_Estimates_PNW.csv"))

ensemble_SOC <- read.csv(here("data/dataframes/ensemble_Frac_Estimates_PNW.csv"))
```

### Tables

Functions for changing tables

```{r}
#| cache: true


std_border <- officer::fp_border(color = "gray10", width = 0.1)
thick_border <- officer::fp_border(color = "black", width = 1.5)
flextable_fun <- function(table){
    table |> flextable() |>
    #colformat_num(j = 7, options(digits=4) ) |>
    flextable::font(fontname = "Aptos", part = "all") |> 
    fontsize(size = 8, part = "body") |>
    fontsize(size = 9, part = "header") |>
    set_table_properties(layout = "autofit", align = "center", width = 1) |>
    flextable::align(align = "center", part = "all") |>
    flextable::align(align = "left", j = 1) |>
    flextable::bg(bg = "gray90", part = "header") |> 
    border_inner_v(part = "header", border = std_border) |> 
    hline(part = "footer", border = thick_border) |> 
    hline(part = "header", border = thick_border) |>
    padding(padding.top = 2, padding.bottom = 2, part = "body") |>
    padding(padding = 5, part = "header") |>
    paginate(init = TRUE, hdr_ftr = TRUE) 
}

mutate_tab <- function(table){
    table |> mutate(Predictor = replace(Predictor, 
                               Predictor == "sd_(Intercept)|sample_ID", 
                               "St.Dev. Random Intercept"),
           Predictor = replace(Predictor, 
                               Predictor == "sd_lower_depth|sample_ID", 
                               "St.Dev. Random Slope and Intercept"),
           Predictor = replace(Predictor, 
                               Predictor == "cor_lower_depth.(Intercept)|sample_ID", 
                               "Random Slope and Intercept Correlation"),
           Predictor = replace(Predictor, 
                               Predictor == "sigma", 
                               "Model Residuals"),
           Predictor = replace(Predictor, 
                               Predictor == "(Intercept)", 
                               "Model Intercept"),
           Predictor = replace(Predictor, 
                               Predictor == "CHM", 
                               "Canopy Height"),
           # Predictor = replace(Predictor, 
           #                     Predictor == "WIP", 
           #                     "Wetland Intrinsic Potential"),
           Predictor = replace(Predictor, 
                               Predictor == "HLI", 
                               "Heat Load Index"),
           Predictor = replace(Predictor, 
                               Predictor == "MAP", 
                               "Mean Annual Precipitation"),
           Predictor = replace(Predictor, 
                               Predictor == "lower_depth", 
                               "Depth"),
           Predictor = str_replace(Predictor, 
                                "GEO", 
                               "Geologic Age: "),
           Predictor = str_replace(Predictor, 
                                   "LITH", 
                                   "PM "),
           Predictor = str_replace(Predictor, 
                                   "PET_MAP", 
                                   "PET:MAP"),
           Predictor = str_replace(Predictor, 
                                   "geomorphons", 
                                   "Landform "))
}



```

WIP Model summary

```{r}
All_WIP_Acc_DF_table <- All_WIP_Acc_DF |>
  mutate(dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) |> 
  dplyr::rename("Study Area" = StudyArea,
                "Overall\nAccuracy" = OverallAcc,
                "Wetland\nOmission Error" = OmissionWet,
                "Wetland\nCommission Error" = ComissionWet) |> 
  flextable_fun()
All_WIP_Acc_DF_table

saveRDS(All_WIP_Acc_DF_table, here("writing/Tables/All_WIP_Acc_DF_table.rds"))

```

WIP Predictors

```{r}
#| cache: true


wipPredictors_table <- wipPredictors |> 
  dplyr::rename("Study Area" = StudyArea) |> 
  mutate(Predictor = case_when(str_detect(Predictor, "dev_") ~ str_replace(Predictor, "dev_", "Deviation "),
                               str_detect(Predictor, "grad_") ~ str_replace(Predictor, "grad_", "Gradient "), 
                               str_detect(Predictor, "plan_") ~ str_replace(Predictor, "plan_", "Planform Curvature  "),
                               str_detect(Predictor, "prof_") ~ str_replace(Predictor, "prof_", "Profile Curvature "),
                               str_detect(Predictor, "twi_") ~ str_replace(Predictor, "twi_", "TWI "),
                               str_detect(Predictor, "HLEF_CHM_") ~ "Canopy Height ",
                               str_detect(Predictor, "HLEF_DEM|DEM") ~ "Elevation  ",
                               str_detect(Predictor, "HLEF_Intensity") ~ str_replace(Predictor, "HLEF_Intensity", "Lidar Intensity  "),
                               str_detect(Predictor, "depthToAny") ~ "Depth to Restrictive Layer",
                               str_detect(Predictor, "Ksat_") ~ "Soil Ksat",
                               str_detect(Predictor, "Hoh_TreeHeight") ~ "Canopy Height",
                               str_detect(Predictor, "NAIP_2017_Hoh_NDVI") ~ "NDVI",
                               str_detect(Predictor, "Band_1") ~ "NDWI",
                               str_detect(Predictor, "Hoh_DTW_allrivers|DTW") ~ "Depth to Water",
                               str_detect(Predictor, "Hoh_all_DTM_NAD83_TWI_4m_v4|TWI") ~ "Topographic Wetness Index",
                               str_detect(Predictor, "CHM") ~ "Canopy Height",
                               str_detect(Predictor, "0\ ") ~ str_replace(Predictor, "0\ ", "0m"),
         .default = Predictor)) |> 
  mutate(Predictor = str_replace(Predictor, "(\\d+)$", "\\1m")) |> 
  mutate(Predictor = str_replace(Predictor, "Hoh_all_DTM_NAD83_TWI_4m_v4|TWI", "Topographic Wetness Index")) |>
  flextable_fun() |> 
  flextable::hline(i = c(24, 43, 61, 78)) |>
  flextable::merge_v(j = 2)
wipPredictors_table

saveRDS(wipPredictors_table, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Tables/wipPredictors_table.rds")
```

SOC predictors 

```{r}
#| cache: true

descriptions <- c(
  Depth      = "Lower depth of sample (cm)",
  WIP              = "Wetland intrinsic potential (probability)",
  MAP              = "Mean annual precipitation (mm)",
  MAT              = "Mean annual temperature (°C)",
  `Parent Material (PM)`            = "Lithology / parent material",
  `PET:MAP`          = "Potential evapotranspiration to precipitation ratio",
  `Canopy Height`              = "Canopy height model (m)",
  HLI              = "Heat load index",
  EVI              = "Enhanced vegetation index",
  NDVI             = "Normalized difference vegetation index",
  Elevation              = "Digital terrain model (elevation, m)",
  Landform      = "Geomorphon landform classification",
  Clay = "Clay content (%)",
  Silt = "Silt content (%)",
  Sand = "Sand content (%)", 
  pH = "Soil pH"
)


pred_ft_data <- tibble(
  Predictor   = names(descriptions),
  Description = descriptions
) |>
  mutate(Source = case_when(Predictor == "Depth" ~ "Field sampling depth interpolated and combined into 25cm intervals. Use as a model predictor is described more in the Modeling Approach section of the methods",
                            Predictor == "WIP" ~ "Derived from modeling with the Wetland Intrisic Potential tool",
                            str_detect(Predictor, "HLI|EVI|NDVI") ~ "Remote sensing index from Sentinel-2, 5 year median of summer growing season 2018 to 2023", 
                            str_detect(Predictor, "Canopy|Elevation|Landform") ~ "Aerial lidar point cloud derived",
                            str_detect(Predictor, "MAP|MAT|PET:MAP") ~ "PRISM climate data for the HRW, MRW, and CVW; WorldClim 2.1 for HLEF",
                            str_detect(Predictor, "Parent") ~ "Surficial geology from WA DNR and USGS",
                            .default = "Field sampling and lab processing"),
         Model = case_when(str_detect(Predictor, "Clay|Silt|Sand|pH") ~ "Mineral SOC %",
                           .default = "SOC stock and Mineral SOC %")) |> 
  dplyr::arrange(desc(Model), Source)

pred_h_i <- which(pred_ft_data$Model != lead(pred_ft_data$Model))
pred_ft <- pred_ft_data |> 
  flextable_fun() |>
  flextable::merge_v(j = 4) |> 
  hline(i = pred_h_i)
pred_ft

saveRDS(pred_ft, file = here("writing/Tables/socPredictors_table.rds"))
```


```{r}
#| cache: true

all_sites_depth_stocks <- pnw_dat |> #group_by(lower_depth) |> 
    summarise(site = "All", 
              lower_depth = 0,
              mean = mean(SOC_stock_spline),
               sd = sd(SOC_stock_spline),
               n = n()) |>
  mutate(se = sd/sqrt(n),
         dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2))))

overall_depth_stocks <- pnw_dat |> group_by(lower_depth) |> 
    summarise(site = "All", 
              mean = mean(SOC_stock_spline),
               sd = sd(SOC_stock_spline),
               n = n()) |>
  mutate(se = sd/sqrt(n),
         dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) |>
  dplyr::select(site, lower_depth, mean, sd, n, se)

site_depth_stocks <- pnw_dat |> group_by(site) |> 
    summarise(lower_depth = 0,
              mean = mean(SOC_stock_spline),
               sd = sd(SOC_stock_spline),
               n = n()) |>
  mutate(se = sd/sqrt(n),
         dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2))))

mean_stdev_stocks_table <- pnw_dat |> group_by(site, lower_depth) |> 
  summarise(mean = mean(SOC_stock_spline),
            sd = sd(SOC_stock_spline),
            n = n()) |>
  mutate(se = sd/sqrt(n),
         dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) |> 
  bind_rows(overall_depth_stocks) |>
  bind_rows(site_depth_stocks) |> 
  bind_rows(all_sites_depth_stocks) |> 
  filter(lower_depth <151) |>
  mutate(site = case_when(str_detect(site, "(?i)HOH") ~ "HRW",
                            str_detect(site, "(?i)MAS") ~ "MRW",
                            str_detect(site, "(?i)COL") ~ "CVW",
                            str_detect(site, "(?i)HLEF") ~ "HLEF",
                            .default = "All"),
         site = factor(site),
         site = fct_relevel(site, c("All", "HLEF", "HRW", "MRW", "CVW"))) |>
    arrange(site) |> 
    rename("Study Area" = site,
           "Depth" = lower_depth) |>
    dplyr::select(-sd) |> 
    relocate(`Study Area`, Depth, mean, se, n) |> 
    flextable_fun() |>
    merge_v(j = ~ `Study Area`) |> 
    mk_par(
    part = "header", j = 1,
    value = as_paragraph("Study\nArea")
    ) |>
    mk_par(
    part = "header", j = 3,
    value = as_paragraph("Mean SOC Stock\n(g cm", as_sup("-2"), ")")
    ) |> 
    mk_par(
    part = "header", j = 4,
    value = as_paragraph("Std. Error SOC Stock\n(g cm", as_sup("-2"), ")")
    ) |> 
    mk_par(
    part = "body", i = c(7,14,21,28,35), j = 2,
    value = as_paragraph("Overall")
    ) |>
    hline(part = "body", border = fp_border_default(width =0)) |> 
    add_body_row(top = FALSE, values = "", colwidths = c( 5)) |>
    hline(part = "body", i = c(7,14,21,28)) |>
    hline(part = "body", i = 35, border = thick_border) |>
    fontsize(size = 10, part = "body", j = 1)

mean_stdev_stocks_table
saveRDS(mean_stdev_stocks_table, file = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Tables/mean_stdev_stocks_table.rds")


```

SOC % by site and depth(?)

```{r}
#| cache: true

percSOC_table <- pnw_dat_lab |> mutate(depth_zone = case_when(lower_depth <= 25 ~ "0-25",
                                             lower_depth > 25 & lower_depth <=50 ~ "25-50",
                                             lower_depth > 50 & lower_depth <=75 ~ "50-75",
                                             lower_depth > 75 & lower_depth <=100 ~ "75-100",
                                             lower_depth > 100 & lower_depth <=125 ~ "100-125",
                                             lower_depth > 125 & lower_depth <=150 ~ "125-150",
                                             .default = " > 150"),
                      wetupl = case_when(WIP >= 0.50 ~ "wet",
                                         .default = "upl")) |> 
  group_by(site, wetupl, .drop = TRUE) |> 
  summarise(mean_percSOC = mean(carbon_perc),
            sd_percSOC = sd(carbon_perc),
            n = n()) |>
  mutate(se_percSOC = sd_percSOC/sqrt(n),
         # depth_zone = ordered(depth_zone, levels = 
         #                                   c("0-25", "25-50","50-75",
         #                                     "75-100","100-125", "125-150", 
         #                                     " > 150")),
         site = case_when(str_detect(site, "(?i)HOH") ~ "HRW",
                            str_detect(site, "(?i)MAS") ~ "MRW",
                            str_detect(site, "(?i)COL") ~ "CVW",
                            str_detect(site, "(?i)HLEF") ~ "HLEF",
                            .default = "All"),
         site = as.factor(site),
         wetupl = case_when(str_detect(wetupl, "wet") ~ "Wetland",
                            .default = "Upland"),
         dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) |> 
  dplyr::select(site, wetupl, mean_percSOC, sd_percSOC, se_percSOC, n) |> 
  mutate(site = fct_relevel(site, c("HLEF", "HRW", "MRW", "CVW"))) |>
  dplyr::arrange(site) |> 
  dplyr::select(-sd_percSOC) |> 
  flextable_fun() |> 
  merge_v(j = c(1)) |> 
  mk_par(
    part = "header", j = 1,
    value = as_paragraph("Study\nArea")
    ) |>
  mk_par(
    part = "header", j = 2,
    value = as_paragraph("Landscape\nClass")
    ) |>
  mk_par(
    part = "header", j = 3,
    value = as_paragraph("Mean\n Mineral SOC %")
    ) |>
  # mk_par(
  #   part = "header", j = 4,
  #   value = as_paragraph("Std. Deviation \nSOC %")
  #   ) |>
  mk_par(
    part = "header", j = 4,
    value = as_paragraph("Std. Error \n Mineral SOC %")
  ) |>
    hline(part = "body", i = 2, j = 2:5) |>
    hline(part = "body", i = 4, j = 2:5) |>
    hline(part = "body", i = 6, j = 2:5) |>
    hline(part = "body", i = 8, border = thick_border) |> 
  fix_border_issues()

percSOC_table
saveRDS(percSOC_table, file = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Tables/SOC_percent_bySite.rds")
```

1m Stocks from samples

```{r}
#| eval: false
full_1m_stocks <- wa_stocks_dat |> 
     mutate(wetupl = "Full") |>
   summarise(site = "Overall",
       wetupl = first(wetupl),
                                mean = mean(SOC_stock_100),
                                sd = sd(SOC_stock_100),
                                n = n()) |>
  mutate(se = sd/sqrt(n),
         dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) 
all_1m_stocks <- wa_stocks_dat |> 
    mutate(wetupl = case_when(WIP >= 0.5 ~ "Wetland",
                              WIP < 0.5 ~  "Upland")) |>
    group_by(wetupl) |> 
    summarise(site = "All",
        wetupl = first(wetupl),
                                mean = mean(SOC_stock_100),
                                sd = sd(SOC_stock_100),
                                n = n()) |>
  mutate(se = sd/sqrt(n),
         dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) |> 
    rbind(full_1m_stocks)


full_sites_1m_stocks <- wa_stocks_dat |> 
     mutate(wetupl = "Full") |>
    group_by(site) |> summarise(wetupl = first(wetupl),
                                mean = mean(SOC_stock_100),
                                sd = sd(SOC_stock_100),
                                n = n()) |>
  mutate(se = sd/sqrt(n),
         dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) 


wa_stocks_dat_table <- wa_stocks_dat |> 
    mutate(wetupl = case_when(WIP >= 0.5 ~ "Wetland",
                              .default = "Upland")) |> 
    group_by(site, wetupl) |> summarise(mean = mean(SOC_stock_100),
                                        sd = sd(SOC_stock_100),
                                        n = n()) |>
  mutate(se = sd/sqrt(n),
         dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) |> 
    rbind(all_1m_stocks) |> 
    rbind(full_sites_1m_stocks) |> 
    mutate(site = case_when(str_detect(site, "HOH") ~ "HRW",
                            str_detect(site, "MAS") ~ "MRW",
                            str_detect(site, "COL") ~ "CVW",
                            .default = "All")) |> 
    mutate(site = factor(site)) |> 
    mutate(site = fct_relevel(site, c("All", "HRW", "MRW", "CVW"))) |>
    arrange(site, wetupl) |> 
    relocate(site, wetupl, mean, sd, se, n) |> 
    rename("Site" = site,
           "Landscape Class" = wetupl) |>
    flextable_fun() |>
    merge_v(j = ~ Site) |> 
    mk_par(
    part = "header", j = 3,
    value = as_paragraph("Mean 1m SOC Stock\n(g cm", as_sup("-2"), ")")
    ) |> 
    mk_par(
    part = "header", j = 4,
    value = as_paragraph("Std. Deviation 1m SOC Stock\n(g cm", as_sup("-2"), ")")
    ) |> 
    mk_par(
    part = "header", j = 5,
    value = as_paragraph("Std. Error 1m SOC Stock\n(g cm", as_sup("-2"), ")")
    ) |> 
    hline(part = "body", border = fp_border_default(width =0)) |> 
    add_body_row(top = FALSE, values = "", colwidths = c( 6)) |>
    hline(part = "body", i = 12, border = thick_border) |> 
    hline(part = "body", i = 3) |>
    hline(part = "body", i = 6) |>
    hline(part = "body", i = 9) |>
    fontsize(size = 10, part = "body", j = 1)
wa_stocks_dat_table
saveRDS(wa_stocks_dat_table, file = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Tables/Stocks_wetup_table.rds")
```

Parial correlation table

```{r}
#| cache: true

library(ppcor)

pnw_lab_wet_ppcor <- pnw_dat_lab |> filter(WIP >= 0.50) |> 
  dplyr::select(carbon_perc, lower_depth, PET_MAP, pH, Clay) |> 
  ppcor::pcor(method = "spearman") 
pnw_lab_upl_ppcor <- pnw_dat_lab |> filter(WIP < 0.50) |> 
  dplyr::select(carbon_perc, lower_depth, PET_MAP, pH, Clay) |> 
  ppcor::pcor(method = "spearman") 
pnw_wet_ppcor <- pnw_dat |> filter(WIP >= 0.50) |> 
  dplyr::select(SOC_stock_spline, lower_depth, PET_MAP, 
                MAT, WIP, CHM, (LITH), (geomorphons)) |> 
  mutate(LITH = rank(LITH), 
         geomorphons = rank(geomorphons)) |> 
  ppcor::pcor(method = "spearman") 
pnw_upl_ppcor <- pnw_dat |> filter(WIP < 0.50) |> 
  dplyr::select(SOC_stock_spline, lower_depth, PET_MAP, 
                MAT, WIP, CHM, (LITH), (geomorphons)) |> 
  mutate(LITH = rank(LITH), 
         geomorphons = rank(geomorphons)) |> 
  ppcor::pcor(method = "spearman") 

ppcor_func <- function(ppcor_df, iv) {
  ppcor_col <- ppcor_df$estimate |> data.frame() |> tibble::rownames_to_column(var = "Predictor") |>  tibble::as_tibble() |> dplyr::select(Predictor, !!iv)
  return(ppcor_col)
}

wetlab_ppcor <- ppcor_func(pnw_lab_wet_ppcor, "carbon_perc")
upllab_ppcor <- ppcor_func(pnw_lab_upl_ppcor, "carbon_perc")
wet_ppcor <- ppcor_func(pnw_wet_ppcor, "SOC_stock_spline")
upl_ppcor <- ppcor_func(pnw_upl_ppcor, "SOC_stock_spline")


all_lab_ppor <- wetlab_ppcor |> rename(carbon_perc_wet = carbon_perc) |> left_join(upllab_ppcor, by = join_by(Predictor)) |> rename(carbon_perc_upl = carbon_perc) |> filter(Predictor != "carbon_perc")

all_ppor <- wet_ppcor |> rename(SOC_stock_spline_wet = SOC_stock_spline) |> left_join(upl_ppcor, by = join_by(Predictor)) |> rename(SOC_stock_spline_upl = SOC_stock_spline) |> filter(Predictor != "SOC_stock_spline")
```

```{r results='asis'}
#| cache: true

all_lab_ppor_ft <- all_lab_ppor |> 
  mutate_tab() |>  
  mutate(dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) |> 
  flextable_fun() |> 
  mk_par(
    part = "header", j = 2,
    value = as_paragraph("Mineral SOC %", "\nWetland (WIP ≥ 50%)")
    ) |> 
  mk_par(
    part = "header", j = 3,
    value = as_paragraph("Mineral SOC %", "\nUpland (WIP < 50%)")
    ) 
all_ppor_ft <- all_ppor |> 
  mutate_tab() |> 
  mutate(dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))) |> 
  flextable_fun() |> 
    mk_par(
    part = "header", j = 2,
    value = as_paragraph("SOC Stock (g cm", as_sup("-2"), ")", "\nWetland (WIP ≥ 50%)")
    ) |> 
  mk_par(
    part = "header", j = 3,
    value = as_paragraph("SOC Stock (g cm", as_sup("-2"), ")", "\nUpland (WIP < 50%)")
    ) 

saveRDS(all_lab_ppor_ft, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Tables/all_lab_ppor_ft.rds")
saveRDS(all_ppor_ft, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Tables/all_ppor_ft.rds")
```

Map table stats

```{r}
#| cache: true

PNW_Mapped_SOC <- PNW_SOC_df |> 
  #bind_rows(gsoc_frac_pnw) |> 
  bind_rows(ensemble_SOC) |> 
  #bind_rows(mcnicol_soc_frac_form) |> 
  mutate(StudyArea = case_when(str_detect(StudyArea, "(?i)HOH") ~ "HRW",
                            str_detect(StudyArea, "(?i)MAS") ~ "MRW",
                            str_detect(StudyArea, "(?i)COL") ~ "CVW",
                            str_detect(StudyArea, "(?i)HLEF") ~ "HLEF",
                            .default = NA)) |> 
  bind_rows(y = nwca_soc_est) |> 
  mutate(Source = case_when(str_detect(Name, "(?i)ensemble") ~ "Ensemble Model",
                            str_detect(Name, "(?i)mcnicol") ~ "McNicol et al., 2019",
                            str_detect(Name, "(?i)nwca") ~ "Uhran et al., 2021",
                            .default = "This Study")) |> 
  mutate(Name = case_when(#str_detect(Name, "(?i)list") ~ "Total",
                            str_detect(Name, "(?i)nwca") ~ "Wetland",
                            str_detect(Name, "(?i)_wet") ~ "Wetland",
                            str_detect(Name, "(?i)_upl") ~ "Upland",
                            .default = "Total")) |> 
  dplyr::arrange(match(Name, c("Total", "Wetland", "Upland")),
                 match(StudyArea, c("HLEF", "HRW", "MRW", "CVW"))) |> 
  dplyr::select( Name, StudyArea, Source, everything()) |>
  dplyr::select(-Perc10thSOC_Mgha, -Perc90thSOC_Mgha) |>
  dplyr::mutate(across(where(is.numeric), ~ case_when(
    . > 100  ~ round(., 0),  # Round to the nearest hundred
    #. > 100 & . <= 1000 ~ round(., 0),  # Round to the nearest ten
    . > 10 & . <= 100 ~ round(., 1),
    . > 0 & . <= 10  ~ round(., 2),  # Round to the nearest whole number
    TRUE ~ .  # Default case (if needed)
  ))) |> 
  flextable_fun() |>
  flextable::compose(part = "header", i=1, j = 1,
               value = c(as_paragraph("Landscape \nClass")  )) |>
  flextable::compose(part = "header", i=1, j = 2,
               value = c(as_paragraph("Study\nArea")  )) |>
  flextable::compose(part = "header", i=1, j = 4,
               value = c(as_paragraph("Total \nArea\n(ha)")  )) |>
  flextable::compose(part = "header", i=1, j = 5,
               value = c(as_paragraph("Mean 1m \nSOC Stock\n", 
                                      "(Mg ha", as_sup("-1"), ")"))) |>
  flextable::compose(part = "header", i=1, j = 6,
               value = c(as_paragraph("Total 1m\nSOC Stock\n(Tg)")  )) |>
  merge_v(j = c(1,2)) |>
  hline(part = "body", i = 8) |>
  hline(part = "body", i = 19) |>
  #hline(part = "body", i = 27) |> 
  hline(part = "body", i = 2, j = 2:6) |> 
  hline(part = "body", i = 4, j = 2:6) |> 
  hline(part = "body", i = 6, j = 2:6) |> 
  hline(part = "body", i = 10, j = 2:6) |> 
  hline(part = "body", i = 13, j = 2:6) |> 
  hline(part = "body", i = 16, j = 2:6) |> 
  hline(part = "body", i = 21, j = 2:6) |> 
  hline(part = "body", i = 23, j = 2:6) |> 
  hline(part = "body", i = 25, j = 2:6) |> 
  #hline(part = "body", i = 28, j = 2:6) |> 
  fix_border_issues()


PNW_Mapped_SOC

saveRDS(PNW_Mapped_SOC, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Tables/PNW_Mapped_SOC.rds")
```

Uncertainty prediction intervals from RFM

```{r}
#| cache: true

mappedSOC_025_1m_depths_sum_df <- read.csv("/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/data/dataframes/PNW_mappedSOC_025_1m_depths_sum_df.csv") |>  mutate(across(contains("SOC"), ~ . * 100))
mappedSOC_975_1m_depths_sum_df <- read.csv("/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/data/dataframes/PNW_mappedSOC_975_1m_depths_sum_df.csv") |>  mutate(across(contains("SOC"), ~ . * 100)) 

uncertainty_table_mapped_SOC <- mappedSOC_025_1m_depths_sum_df |> 
  dplyr::select(-Perc10thSOC_Mgha, -Perc90thSOC_Mgha) |> 
  rename(Perc025_AverageSOC_Mgha = AverageSOC_Mgha,
         Perc025_Total_SOC_Tg = Total_SOC_Tg) |> 
  cbind(mappedSOC_975_1m_depths_sum_df[,c("AverageSOC_Mgha", "Total_SOC_Tg")]) |> 
  rename(Perc975_AverageSOC_Mgha = AverageSOC_Mgha,
         Perc975_Total_SOC_Tg = Total_SOC_Tg)  |> 
  mutate(StudyArea = case_when(str_detect(StudyArea, "(?i)HOH") ~ "HRW",
                            str_detect(StudyArea, "(?i)MAS") ~ "MRW",
                            str_detect(StudyArea, "(?i)COL") ~ "CVW",
                            str_detect(StudyArea, "(?i)HLEF") ~ "HLEF",
                            .default = NA)) |> 
  mutate(Name = case_when(#str_detect(Name, "(?i)list") ~ "Total",
                            str_detect(Name, "(?i)_wet") ~ "Wetland",
                            str_detect(Name, "(?i)_upl") ~ "Upland",
                            .default = "Total")) |> 
  dplyr::arrange(match(Name, c("Total", "Wetland", "Upland")),
                 match(StudyArea, c("HLEF", "HRW", "MRW", "CVW"))) |> 
  dplyr::mutate(across(where(is.numeric), ~ case_when(
    . > 100  ~ round(., 0),  # Round to the nearest hundred
    #. > 100 & . <= 1000 ~ round(., 0),  # Round to the nearest ten
    . > 10 & . <= 100 ~ round(., 1),
    . > 0 & . <= 10  ~ round(., 2),  # Round to the nearest whole number
    TRUE ~ .  # Default case (if needed)
  ))) |> 
  mutate(Perc95th_AverageSOC_Mgha = paste0(Perc025_AverageSOC_Mgha, " - ", Perc975_AverageSOC_Mgha),
         Perc95th_TotalSOC_Tg = paste0(Perc025_Total_SOC_Tg, " - ", Perc975_Total_SOC_Tg)) |> 
  dplyr::select(Name, StudyArea, Total_area, Perc95th_AverageSOC_Mgha, Perc95th_TotalSOC_Tg) |> 
  flextable_fun() |>
  flextable::compose(part = "header", i=1, j = 1,
               value = c(as_paragraph("Landscape \nClass")  )) |>
  flextable::compose(part = "header", i=1, j = 2,
               value = c(as_paragraph("Study\nArea")  )) |>
  flextable::compose(part = "header", i=1, j = 3,
               value = c(as_paragraph("Total \nArea (ha)")  )) |>
  flextable::compose(part = "header", i=1, j = 4,
               value = c(as_paragraph("95% Prediction Interval\nMean 1m SOC Stock\n", 
                                      ("Mg ha"), as_sup("-1")))) |>
  flextable::compose(part = "header", i=1, j = 5,
               value = c(as_paragraph("95% Prediction Interval\nTotal 1m SOC Stock \n(Tg)")  )) |>
  merge_v(j = c(1,2)) |> 
  hline(part = "body", i = 4, j = 2:5) |> 
  hline(part = "body", i = 8, j = 2:5) |> 
  fix_border_issues()
uncertainty_table_mapped_SOC

saveRDS(uncertainty_table_mapped_SOC, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Tables/PNW_Uncertainty_Mapped_SOC.rds")
```

NWI wetland areas - use wetland area estimates from the mapped data tables

```{r}
#| cache: true

NWI_areas_df <- read.csv("/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/data/dataframes/PNW_NWI_wetland_areas.csv")

PNW_Wet_Area <- PNW_SOC_df |> dplyr::filter(Name == "list_wet") |> dplyr::select(StudyArea, Total_area) |> 
  dplyr::mutate(StudyArea = str_replace_all(StudyArea, c(hlef_soc = "HLEF",
                hoh_soc = "HRW",
                mas_soc = "MRW",
                col_soc = "CVW"))) |> 
  left_join(NWI_areas_df, by = join_by(StudyArea)) |> 
  dplyr::mutate(Perc_change = ((Total_area - NWI_wet_ha)/NWI_wet_ha)*100,
                across(where(is.numeric), ~ case_when(
                        . > 100  ~ round(., 0),  # Round to the nearest hundred
                        . > 100 & . <= 1000 ~ round(., 0),  # Round to the nearest ten
                        . > 10 & . <= 100 ~ round(., 1),
                        . > 0 & . <= 10  ~ round(., 2),  # Round to the nearest whole number
                        TRUE ~ .  # Default case (if needed)
  )))

PNW_Wet_Area_Table <- PNW_Wet_Area |> 
  #dplyr::mutate(Perc_change = paste0(Perc_change, " %")) |> 
  flextable_fun() |>
  flextable::compose(part = "header", i=1, j = 1,
               value = c(as_paragraph("Study\nArea"))) |>
  flextable::compose(part = "header", i=1, j = 2,
               value = c(as_paragraph("WIP Wetland\nArea (ha)")  )) |>
  flextable::compose(part = "header", i=1, j = 3,
               value = c(as_paragraph("NWI Wetland\nArea (ha)")  )) |>
  flextable::compose(part = "header", i=1, j = 4,
               value = c(as_paragraph("% Change\n(WIP-NWI)/NWI")  )) 
PNW_Wet_Area_Table
saveRDS(PNW_Wet_Area_Table, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Tables/PNW_Wet_Area_Table.rds")
```

### Model Performance Table

Test dataset metrics only

```{r}
#| cache: true

stock_mods_table <- mod_errdf |> 
    dplyr::select(LandscapeClass, LMM.LogMargR2, LMM.LogCondR2,
                  LMM.nLogPredObsR2, LMM.nlogPredObsRMSE) |> 
    left_join(y = qrf_errdf |> 
                  dplyr::select(LandscapeClass, RFM.R2_OOB, RFM.RMSE_OOB,
                                RFM.R2_test, RFM.RMSE_test), 
              by = join_by(LandscapeClass)) |> 
    dplyr::mutate(`Model Prediction` = "SOC Stock") |> 
    dplyr::select(`Model Prediction`, everything()) |> 
    dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))
stock_mods_table_noWIP <- mod_errdf_noWIP |> 
    dplyr::select(LandscapeClass, LMM.LogMargR2, LMM.LogCondR2,
                  LMM.nLogPredObsR2, LMM.nlogPredObsRMSE) |> 
    left_join(y = qrf_errdf_noWIP |> 
                  dplyr::select(LandscapeClass, RFM.R2_OOB, RFM.RMSE_OOB,
                                RFM.R2_test, RFM.RMSE_test), 
              by = join_by(LandscapeClass)) |> 
    dplyr::mutate(`Model Prediction` = "SOC Stock NW") |> 
    dplyr::select(`Model Prediction`, everything()) |> 
    dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))

Cperc_mods_table <- mod_errdf_lab |> 
    dplyr::select(LandscapeClass, LMM.LogMargR2, LMM.LogCondR2,
                  LMM.nLogPredObsR2, LMM.nlogPredObsRMSE) |>
    left_join(y = qrf_errdf_lab|> 
                  dplyr::select(LandscapeClass, RFM.R2_OOB, RFM.RMSE_OOB,
                                RFM.R2_test, RFM.RMSE_test), 
              by = join_by(LandscapeClass)) |>
    dplyr::mutate(`Model Prediction` = "SOC Perc") |> 
    dplyr::select(`Model Prediction`, everything()) |> 
    dplyr::mutate(across(where(is.numeric), ~base::round(.x, 2)))
```


```{r}
#| cache: true

stat_table_format <- stock_mods_table |> 
    rbind(stock_mods_table_noWIP, 
          Cperc_mods_table)  |>
    rename_with(~gsub("LMM", "Linear Mixed",.)) |> 
    rename_with(~gsub("RFM", "Random Forest",.)) |> 
    flextable_fun() |> 
    merge_v(j = 1) |> 
    separate_header(split = "[\\.]", fixed = F) |>
    # flextable::compose(part = "body", i=c(2:3,5:6, 8:9), j = c(3:4, 7:8),
    #            value = c(as_paragraph("---")  )) |>
    align(align = "center", part =  "header") |>
    vline(part = "header",  j = 6) |>
    flextable::compose(part = "header", i=1, j = 2,
               value = c(as_paragraph("Landscape \nClass")  )) |>
    flextable::compose(part = "header", i = 2, j = 3,
               value = c(as_paragraph("Marginal R", as_sup("2"),
                                      "\n"))) |>
    flextable::compose(part = "header", i = 2, j = 4,
               value = c(as_paragraph("Conditional R", as_sup("2"),
                                      "\n"))) |>
    flextable::compose(part = "header", i = 2, j = 5,
               value = c(as_paragraph("Test R", as_sup("2")))) |>
    flextable::compose(part = "header", i = 2, j = 6,
               value = c(as_paragraph("Test RMSE", (" \ng cm"), as_sup("-2")))) |>
    flextable::compose(part = "header", i = 2, j = 7,
               value = c(as_paragraph("OOB R", as_sup("2"),
                                      "\n"))) |>
    flextable::compose(part = "header", i = 2, j = 8,
               value = c(as_paragraph("OOB RMSE", (" \ng cm"), as_sup("-2")))) |>
    flextable::compose(part = "header", i = 2, j = 9,
           value = c(as_paragraph("Test R", as_sup("2")))) |>
    flextable::compose(part = "header", i = 2, j = 10,
               value = c(as_paragraph("Test RMSE", (" \ng cm"), as_sup("-2")))) |>

    flextable::compose(part = "body", i = 1, j = 1,
               value = c(as_paragraph("SOC stock", 
                                      "\ng cm",
                                      as_sup("-2")))) |>
    flextable::compose(part = "body", i = 4, j = 1,
               value = c(as_paragraph("SOC stock", 
                                      as_sub("No WIP"), 
                                      "\ng cm",
                                      as_sup("-2")))) |>
    flextable::compose(part = "body", i = 7, j = 1,
               value = c(as_paragraph("Mineral SOC \n%"))) |>
    hline(i = c(3, 6), j = 2:10) |> 
    fix_border_issues() 
    # vline(part = "header",  j = 9) |> 
    # delete_rows(i = c(2, 3), part = "body") |> 
    # delete_columns(j = 1) |> 
    # flextable::merge_h_range(i = 1, j1 = 1, j2 = 5, part = "header") |> 
    # flextable::merge_h_range(i = 1, j1 = 6, j2 = 8, part = "header") |> 
    # flextable::merge_h_range(i = 1, j1 = 9, j2 = 11, part = "header") |> 

stat_table_format

saveRDS(stat_table_format, file = here("writing/Tables/Model_stats_table.rds"))
readRDS(file = here("writing/Tables/Model_stats_table.rds"))
```

