---
title: "Figure Formatting"
format: html
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.show = "hold", 
                      time_it = TRUE, dpi = 100)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = T, collapse = TRUE)

library(terra)
library(lme4)
library(MASS)
library(lmerTest)
library(MuMIn)
library(ggplot2)
library(ggeffects)
library(ggtext)
library(merTools)
library(glmnet)
library(stats)
library(ggcorrplot)
library(forcats)
library(RColorBrewer)
library(webshot)
library(kableExtra)
library(tidyverse)
library(flextable)
library(ggnewscale)
library(tidyterra)
library(ggpubr)
library(ggspatial)
library(patchwork)
library(randomForest)
library(mlr3)
library(mlr3pipelines)
library(pdp)
library(kernelshap)
library(shapviz)
library(here)
here()
#knitr::knit_hooks$set(webgl = hook_webgl)
rgl::setupKnitr(autoprint = TRUE)

setGDALconfig("GDAL_PAM_ENABLED", "FALSE")

#tmap_options(raster.max.cells = c(plot = 1e7, view = 1e6))
```

Data sources
```{r} 
#| include: false
#| echo: false
#| message: false
#| warning: false
#| cache: true
# knitr::purl("/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/analysis/WA_SOC_controls/PNW_SOC_Prediction.qmd", output = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/analysis/WA_SOC_controls/PNW_SOC_Prediction.R")

source(here("analysis/All_WA_Functions.R"))
 
```

### Models

```{r}

lmm_stocks <- readRDS(here("analysis/models/PNW_Model_Dredge_Spline.rds"))
lmm_stocks_noWIP <- readRDS(here("analysis/models/PNW_Model_Dredge_Spline_noWIP.rds"))
lmm_lab <- readRDS(here("analysis/models/PNW_Model_Lab_dredge_best.rds"))

QRFM_stocks <- readRDS(here("analysis/models/MLR3_RFM_Ranger_tuned.rds"))
QRFM_stocks_noWIP <- readRDS(here("analysis/models/MLR3_RFM_Ranger_tuned_noWIP.rds"))
QRFM_lab <- readRDS(here("analysis/models/MLR3_RFM_Ranger_tuned_LabData.rds"))

```

### Dataframes

```{r}
All_WIP_Acc_DF <- read.csv(here("data/dataframes/All_WIP_Acc_DF.csv"))
wipPredictors <- read.csv(here("data/dataframes/PNW_WIP_Predictors.csv"))
allsite_WIPimp <- read.csv(here("data/dataframes/PNW_WIP_VarImp.csv"))

pnw_dat <- read.csv(here("data/dataframes/PNW_data_MLR3_RFM_LMM.csv"))
wa_stocks_dat <- read.csv(here("data/dataframes/All_WA_1m_stocks_preds.csv"))

train <- read.csv(here("data/dataframes/Train_data.csv")) |> 
  dplyr::select(-X) |> dplyr::mutate(across(where(is.character), as.factor))
test <- read.csv(here("data/dataframes/Test_data.csv")) |> 
  dplyr::select(-X) |> dplyr::mutate(across(where(is.character), as.factor)) |> 
  mutate(geomorphons = factor(geomorphons, levels = levels(train$geomorphons)))
test_viz <- read.csv(here("data/dataframes/Test_data.csv")) |> 
  dplyr::select( -X) |> dplyr::mutate(across(where(is.character), as.factor)) |> 
  mutate(geomorphons = factor(geomorphons, levels = levels(train$geomorphons)))
test.wet <- read.csv(here("data/dataframes/RFM_Test_Wet_data.csv")) |> 
  dplyr::select(-X) |> dplyr::mutate(across(where(is.character), as.factor)) |> 
  mutate(geomorphons = factor(geomorphons, levels = levels(train$geomorphons)))
test.upl <- read.csv(here("data/dataframes/RFM_Test_Upl_data.csv")) |> 
  dplyr::select(-X) |> dplyr::mutate(across(where(is.character), as.factor)) |> 
  mutate(geomorphons = factor(geomorphons, levels = levels(train$geomorphons)))
RF_testpred <- read.csv(here("data/dataframes/PNW_SOC_RF_test_pred_response.csv")) |> 
  select(-WIP) |> 
  cbind(test)
LMM_testpred <- read.csv(here("data/dataframes/PNW_SOC_LMM_test_pred_response.csv")) |> 
  cbind(test)
RF_testprednw <- read.csv(here("data/dataframes/PNW_SOC_RF_noWIP_test_pred_response.csv")) |> 
  cbind(test)
LMM_testprednw <- read.csv(here("data/dataframes/PNW_SOC_LMM_noWIP_test_pred_response.csv")) |> 
  cbind(test)

pnw_dat_lab <- read.csv(here("data/dataframes/PNW_Lab_data_MLR3_RFM_LMM.csv")) |> 
  dplyr::select( -X)
trainlab<- read.csv(
          file = here("data/dataframes/Train_Lab_data.csv")) |> 
  dplyr::select(-X)
testlab<- read.csv(
          file = here("data/dataframes/Test_Lab_data.csv"))|> 
  dplyr::select(-X)
testlab_viz<- read.csv(
          file = here("data/dataframes/Test_Lab_data.csv"))|> 
  dplyr::select( -X)
testlab_wet<- read.csv(
          file = here("data/dataframes/RFM_Test_Lab_Wet_data.csv"))|> 
  dplyr::select(X)
testlab_upl<- read.csv(
          file = here("data/dataframes/RFM_Test_Lab_Upl_data.csv")) |> 
  dplyr::select(-X)
RF_testlabpred <- read.csv(here("data/dataframes/PNW_SOC_Edaphic_RF_test_pred_response.csv")) |>
  select(-WIP) |> 
  cbind(testlab)
LMM_testlabpred <- read.csv(here("data/dataframes/PNW_SOC_Edaphic_LMM_test_pred_response.csv")) |>
  cbind(testlab)

train_dat_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Train_Scale.csv"))
train_dat_wet_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Train_Wet_Scale.csv"))
train_dat_upl_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Train_Upl_Scale.csv"))
test_dat_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Test_Scale.csv"))
test_dat_wet_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Test_Wet_Scale.csv"))
test_dat_upl_scale <- read.csv(here("data/dataframes/PNW_Spline_Stocks_Test_Upl_Scale.csv"))

trainlab_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Train_Scale.csv"))
trainlab_wet_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Train_Wet_Scale.csv"))
trainlab_upl_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Train_Upl_Scale.csv"))
testlab_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Test_Scale.csv"))
testlab_wet_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Test_Wet_Scale.csv"))
testlab_upl_scale <- read.csv(here("data/dataframes/PNW_Spline_Lab_Test_Upl_Scale.csv"))

anova_ci_stocks <- read.csv(paste0(here("data/dataframes/anova_best_dredge_model_confint95.csv")))
anova_ci_stocks_nw <- read.csv(paste0(here("data/dataframes/anova_best_dredge_model_confint95_noWIP.csv")))
anova_ci_lab <- read.csv(paste0(here("data/dataframes/LMM_Lab_confint95.csv")))


PI_RMSE_stocks <- readRDS(paste0(here("data/dataframes/LMM_PredictInterval_RMSE_Stocks.rds")))
PIw_RMSE_stocks <- readRDS(paste0(here("data/dataframes/LMM_Wet_PredictInterval_RMSE_Stocks.rds")))
PIu_RMSE_stocks <- readRDS(paste0(here("data/dataframes/LMM_PredictInterval_Upl_RMSE_Stocks.rds")))

PI_RMSE_lab <- readRDS(paste0(here("data/dataframes/LMM_PredictInterval_RMSE_Lab.rds")))
PIw_RMSE_lab <- readRDS(paste0(here("data/dataframes/LMM_Wet_PredictInterval_RMSE_Lab.rds")))
PIu_RMSE_lab <- readRDS(paste0(here("data/dataframes/LMM_PredictInterval_Upl_RMSE_Lab.rds")))

mod_errdf <- read.csv(here("data/dataframes/PNW_Stocks_model_fit.csv"))
mod_errdf_noWIP <- read.csv(here("data/dataframes/PNW_Stocks_model_fit_noWIP.csv"))
mod_errdf_lab <- read.csv(here("data/dataframes/PNW_Lab_model_fit.csv"))
qrf_errdf <- read.csv(here("data/dataframes/MLR3_RFM_ErrorDF.csv"))
qrf_errdf_noWIP <- read.csv(here("data/dataframes/MLR3_RFM_ErrorDF_noWIP.csv"))
qrf_errdf_lab <- read.csv(here("data/dataframes/MLR3_RFM_ErrorDF_LabData.csv"))


ks <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap.rds")) 
ks_wet <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_Wet.rds")) 
ks_upl <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_Upl.rds")) 
ks_nw <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_noWIP.rds"))
ks_full <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_Full.rds"))

ia_1 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Interact_1_plot.rds")) 
ia_1$data <- ia_1$data |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_2 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_2_plot.rds")) 
ia_2$results <- ia_2$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_3 <- readRDS( 
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_3_plot.rds")) 
ia_3$results <- ia_3$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_4 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_4_plot.rds")) 
ia_4$results <- ia_4$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_5 <- readRDS( 
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_5_plot.rds"))  
ia_5$results <- ia_5$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)

ks_lab <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_LabData.rds")) 
ks_lab_wet <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_Wet_LabData.rds")) 
ks_lab_upl <- readRDS(file = here("analysis/models/MLR3_RFM_Ranger_KernelShap_Upl_LabData.rds")) 

ia_lab_1 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Interact_Lab_1_plot.rds")) 
ia_lab_1$data <- ia_lab_1$data |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_lab_2 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_Lab_2_plot.rds")) 
ia_lab_2$results <- ia_lab_2$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_lab_3 <- readRDS( 
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_Lab_3_plot.rds")) 
ia_lab_3$results <- ia_lab_3$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_lab_4 <- readRDS(
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_Lab_4_plot.rds")) 
ia_lab_4$results <- ia_lab_4$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)
ia_lab_5 <- readRDS( 
        file = here("writing/Figures/MLR3_RFM_Ranger_Interact_Lab_5_plot.rds")) 
ia_lab_5$results <- ia_lab_5$results |> dplyr::arrange(desc(.interaction)) |> slice(1:5)

```

### Graphs

```{r graph params}

theme_param <- theme(
    text = element_text(size = 8),
    legend.position = 'right', 
    legend.key.size = unit(0.6, "cm"),
    legend.spacing.x = unit(1.2, "cm"),
    legend.text = element_text(size = rel(0.70), hjust = 0.5),
    panel.background = element_blank(),
    panel.grid.major = element_line(colour = "grey80", linewidth = 0.2),
    axis.ticks = element_blank(),
    axis.text = element_text(size = rel(0.8)),
    axis.title = element_text(size = rel(1))
    )

```


```{r}

est_func <- function(data, coef_string){
  if(str_detect(deparse(substitute(data)), "stocks")){
    dat_type <- "stocks"  
  } else {
    dat_type <- "lab"
  }
  coef <- data[data$Predictor == coef_string, "Coefficient"]
  est_2.5 <- data[data$Predictor == coef_string, "conf2.5"]
  est_97.5 <- data[data$Predictor == coef_string, "conf.975"]
  
  est_string <- paste0("scaled coefficient estimate = ", 
                      coef, " 95% CI: ",
                      est_2.5, " - ", est_97.5)
  print(est_string)
  saveRDS(est_string, paste0("/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/in_text_values/",
                             coef_string, "_", dat_type, "_est.rds"))
}
for(i in anova_ci_stocks$Predictor){
  est_func(anova_ci_stocks, i)
  #print(i)
}
for(i in anova_ci_lab$Predictor){
  est_func(anova_ci_lab, i)
  #print(i)
}


```

#### WIP Model Var Imp

```{r}
PNW_WIPimp_plot <- ggplot(allsite_WIPimp, aes(x = Predictor, y = MeanDecreaseAccuracy)) + 
    geom_col() +
    facet_wrap(~ StudyArea) +
    coord_flip() + 
    ylab("Mean Decrease Accuracy") +
    theme_param

PNW_WIPimp_plot

saveRDS(PNW_WIPimp_plot, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/PNW_WIPimp_plot.rds")
```



#### Model fits vs. Predictors



#### Prediction vs. Actual Plots

```{r}

predict_plot <- function(model_testpred, fill_var = NULL) {

  fit_vals  <- model_testpred["predict"]
  true_vals <- model_testpred["truth"]
  axis_max  <- c(max(fit_vals) * 1.1, max(true_vals) * 1.1)[which.max(c(max(fit_vals), max(true_vals)))]
  mod_string <- tolower(deparse(substitute(model_testpred)))

  # WIP values still resolved (used as default fill)
  if (str_detect(mod_string, "lmm") & str_detect(mod_string, "lab")) {
    WIP_vals <- testlab["WIP"]
  } else if (str_detect(mod_string, "lmm") & !str_detect(mod_string, "lab")) {
    WIP_vals <- test["WIP"]
  } else if (str_detect(mod_string, "rf") & str_detect(mod_string, "lab")) {
    WIP_vals <- testlab["WIP"]
  } else if (str_detect(mod_string, "rf") & !str_detect(mod_string, "lab")) {
    WIP_vals <- test["WIP"]
  }

  if (any(str_detect(names(model_testpred), "carbon_perc"))) {
    xlabel <- "Observed Mineral SOC (%)"
    ylabel <- "Predicted Mineral SOC (%)"
  } else if (any(str_detect(names(model_testpred), "SOC_stock_spline"))) {
    xlabel <- expression(Observed ~ SOC ~ g ~ cm^{"-2"})
    ylabel <- expression(Predicted ~ SOC ~ g ~ cm^{"-2"})
  }

  # Resolve fill source: default to WIP, else pull the named column
  use_default_fill <- is.null(fill_var)
  fill_data <- if (use_default_fill) WIP_vals[[1]] else model_testpred[[fill_var]]

  # Graphing
  predict_graph <- ggplot(data = model_testpred) +
    geom_point(aes(y = fit_vals[[1]],
                   x = true_vals[[1]],
                   fill = fill_data,
                   shape = site),
               size = 3.5) +
    geom_abline(intercept = 0, slope = 1) +
    scale_shape_manual(name = "Study\nArea",
                       values = c(21, 22, 24, 25),
                       labels = c("Col" = "CVW", "HLEF" = "HLEF",
                                  "Hoh" = "HRW", "Mas" = "MRW")) +
    ylab(ylabel) +
    xlab(xlabel) +
    xlim(0, axis_max) +
    ylim(0, axis_max) +
    theme_param

  # Only attach the WIP scale when using the default fill,
  # so an overriding scale_fill_* added later won't conflict.
  if (use_default_fill) {
    predict_graph <- predict_graph +
      scale_fill_gradientn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlGnBu"),
                           values  = c(-0.1, 1.1),
                           breaks  = c(0.01, 0.5, 0.99),
                           labels  = c("0%", "50%", "100%"),
                           name    = "Wetland \nIntrinsic \nPotential")
  }

  return(predict_graph)
}

predict_plot(RF_testlabpred)

lmm_lab_r2 <- format(round(mod_errdf_lab[1, "LMM.nLogPredObsR2"], 2), nsmall = 2)
qrf_lab_r2 <- format(round(qrf_errdf_lab[1, "RFM.R2_test"], 2) , nsmall = 2)
lmm_stocks_r2 <- format(round(mod_errdf[1, "LMM.nLogPredObsR2"], 2), nsmall = 2)
lmm_stocks_noWIP_r2 <- format(round(mod_errdf_noWIP[1, "LMM.nLogPredObsR2"], 2), nsmall = 2)
qrf_stocks_r2 <- format(round(qrf_errdf[1, "RFM.R2_test"], digits = 2), nsmall = 2)
qrf_stocks_noWIP_r2 <- format(round(qrf_errdf_noWIP[1, "RFM.R2_test"], digits = 2), nsmall = 2)

```


```{r}
#| label: pred_act_plots
#| fig-width: 6.5
#| fig-height: 5
lmm_lab_plot <- predict_plot(LMM_testlabpred, fill_var = "pH") +
    scale_fill_gradientn(
    colours = RColorBrewer::brewer.pal(n = 9, name = "Reds"),
    breaks  = c(4, 5.5, 7),
    limits  = c(4, 7),
    name    = "pH"
  ) +
    annotate(geom = "text", x = 3, y = 8.2,
             label = as.expression(bquote(R^2 == .(lmm_lab_r2)))) 

lmm_stocks_plot <- predict_plot(LMM_testpred, fill_var = "geomorphons") +
  scale_fill_brewer(
    palette = "Spectral",
    name    = "Landform"
  ) +  guides(fill = guide_legend(override.aes = list(shape = 21))) +
    annotate(geom = "text", x = 1.3, y = 1.9,
             label = as.expression(bquote(R^2 == .(lmm_stocks_r2)))) + 
    annotate("rect", xmin = 0.8, xmax = 1.8, ymin = 0.7, ymax = 1.7, alpha = 0.25, 
             fill = "red")
lmm_stocks_nw_plot <- predict_plot(LMM_testprednw, fill_var = "PET_MAP") +
  scale_fill_gradientn(
    colours = rev(RColorBrewer::brewer.pal(n = 9, name = "BrBG")),
    breaks  = c(0, 1, 2),
    limits  = c(0, 2),
    name    = "PET:MAP"
  ) + 
    annotate(geom = "text", x = 1.3, y = 1.9,
             label = as.expression(bquote(R^2 == .(lmm_stocks_noWIP_r2)))) + 
    annotate("rect", xmin = 0.8, xmax = 1.8, ymin = 0.7, ymax = 1.7, alpha = 0.25, 
             fill = "red")
qrf_stocks_plot <- predict_plot(RF_testpred) +
    annotate(geom = "text", x = 1.3, y = 1.9,
             label = as.expression(bquote(R^2 == .(qrf_stocks_r2)))) +
    annotate("rect", xmin = 0.8, xmax = 1.8, ymin = 0.7, ymax = 1.7, alpha = 0.25, 
             fill = "red")
qrf_stocks_nw_plot <- predict_plot(RF_testprednw) +
    annotate(geom = "text", x = 1.3, y = 1.9,
             label = as.expression(bquote(R^2 == .(qrf_stocks_noWIP_r2)))) + 
    annotate("rect", xmin = 0.8, xmax = 1.8, ymin = 0.7, ymax = 1.7, alpha = 0.25, 
             fill = "red")
qrf_lab_plot <- predict_plot(RF_testlabpred, fill_var = "Clay") +
    scale_fill_gradientn(
    colours = RColorBrewer::brewer.pal(n = 9, name = "Greens"),
    breaks  = c(0, 15, 30),
    limits  = c(0, 30),
    name    = "Clay"
  ) +
    annotate(geom = "text", x = 3, y = 8.2,
             label = as.expression(bquote(R^2 == .(qrf_lab_r2)))) 

# wip_legend <- ggpubr::get_legend(qrf_lab_plot + theme(legend.box = "horizontal",
#                                                       legend.key.width = rel(0.7)), position = "top") 
# as_ggplot(wip_legend) 

design <- "AB\nCD\nEF\nGG\nGG"

model_fit_plots <-
  (lmm_stocks_plot + ggtitle("Linear Mixed")) +
  (qrf_stocks_plot + ggtitle("Random Forest")) +
  lmm_stocks_nw_plot + qrf_stocks_nw_plot +
  lmm_lab_plot + qrf_lab_plot +
  guide_area() +
  patchwork::plot_layout(design = design, axis_titles = "collect",
                         guides = "collect",
                         heights = c(1, 1, 1, 0.3, 0.3)) +
  plot_annotation(tag_levels = list(c("A ", "B", "C", "D", "E", "F", "")),
                  tag_suffix = ".") &
  theme(plot.tag = element_text(size = 8, face = "bold"),
        plot.tag.position = c(0.1, 1.05),
        plot.title = element_text(hjust = 0.5),
        legend.key.width  = unit(0.3, "cm"),
        legend.key.height = unit(0.3, "cm"),
        legend.title = element_text(size = 7),
        legend.text  = element_text(size = 6),
        legend.direction = "horizontal",
        legend.box = "horizontal")
model_fit_plots
saveRDS(model_fit_plots, file = here("writing/Figures/Model_Test_Fit_Scatter.rds"))
```

```{r}
library(ggpubr)

plots_legendless <- list(
  lmm_stocks_plot    + ggtitle("Linear Mixed") + theme(legend.position = "none"),
  qrf_stocks_plot    + ggtitle("Random Forest")  + theme(legend.position = "none"),
  lmm_stocks_nw_plot + theme(legend.position = "none"),
  qrf_stocks_nw_plot + theme(legend.position = "none"),
  lmm_lab_plot       + theme(legend.position = "none"),
  qrf_lab_plot       + theme(legend.position = "none")
)

leg_theme <- theme(legend.direction = "horizontal",
                   legend.title = element_text(size = 7),
                   legend.text  = element_text(size = 6),
                   legend.key.width  = unit(0.3, "cm"),
                   legend.key.height = unit(0.3, "cm"),
                   legend.background     = element_blank(),   # outer box
                   legend.box.background = element_blank(),   # box around grouped legends
                   legend.key            = element_blank())   # per-key background

cbar <- guides(fill = guide_colourbar(title.position = "left",
                                      barwidth  = unit(2, "cm"),
                                      barheight = unit(0.2, "cm")))

# --- Shape legend: extract ONCE, and suppress fill so only shape shows ---
leg_shape <- ggpubr::get_legend(
  qrf_stocks_plot + leg_theme +
    guides(shape = guide_legend(nrow = 1, title.position = "left"),
           fill  = "none")          # drop fill so only Study Area shape remains
)

# --- Each fill legend: extract once, and suppress shape so it doesn't duplicate ---
no_shape <- guides(shape = "none")

leg_landform <- ggpubr::get_legend(lmm_stocks_plot + leg_theme + no_shape +
                                     guides(fill = guide_legend(nrow = 2,
                                                                title.position = "left",
                                            override.aes = list(shape = 21))))
leg_wip      <- ggpubr::get_legend(qrf_stocks_plot + leg_theme + no_shape + cbar)
leg_petmap   <- ggpubr::get_legend(lmm_stocks_nw_plot + leg_theme + no_shape + cbar)
leg_ph       <- ggpubr::get_legend(lmm_lab_plot + leg_theme + no_shape + cbar)
leg_clay     <- ggpubr::get_legend(qrf_lab_plot + leg_theme + no_shape + cbar)

# Row 1: shape + the two discrete-ish / first gradients
legend_row1 <- ggpubr::ggarrange(
  ggpubr::as_ggplot(leg_shape),
  ggpubr::as_ggplot(leg_landform),
  nrow = 1, widths = c(1, 1.3)
)
# Row 2: remaining gradients
legend_row2 <- ggpubr::ggarrange(
  ggpubr::as_ggplot(leg_petmap),
  ggpubr::as_ggplot(leg_ph),
  ggpubr::as_ggplot(leg_clay),
  ggpubr::as_ggplot(leg_wip),
  nrow = 1, widths = c(1, 1, 1, 1)
)
```


```{r}
#| fig-width: 6.76
#| fig-height: 10.36
#| out-width: 6.5in
#| fig-dpi: 300
design <- "
AB
CD
EF
GG
HH
"

model_fit_plots2 <- wrap_plots(plots_legendless) +
  wrap_elements(full = legend_row1) +
  wrap_elements(full = legend_row2) +
  plot_layout(design = design,
              axis_titles = "collect",
              heights = c(1, 1, 1, 0.2, 0.1)) +
  plot_annotation(tag_levels = list(c("A ", "B", "C", "D", "E", "F", "", "")),
                  tag_suffix = ".") &
  theme(plot.tag = element_text(size = 8, face = "bold"),
        plot.tag.position = c(0.1, 1.03),
        plot.title = element_text(hjust = 0.5),
        legend.box = element_blank())

model_fit_plots2
saveRDS(model_fit_plots2, file = here("writing/Figures/Model_Test_Fit_Scatter2.rds"))
```



### Coefficent plots 

```{r}
coef_fig_func <- function(anova_ci, model) {
  if(str_detect(deparse(substitute(model)), "stocks")){
            anova_df <-  anova_ci |> mutate(Predictor = case_when(
              str_detect(Predictor, "Intercept") ~ str_replace(Predictor,
                                                               "\\(Intercept\\)",
                                                 "PM glacial outwash"),
                                                .default = (Predictor)
                    )
              )
  } else{
     anova_df <-  anova_ci |> mutate(Predictor = case_when(
              str_detect(Predictor, "Intercept") ~ str_replace(Predictor,
                                                               "\\(Intercept\\)",
                                                 "Intercept"),
                                                .default = (Predictor)
                    )
              )
  }
  coef_fig <- anova_df |> 
    dplyr::filter(Predictor != "Model" & Predictor != "LITH" & Predictor != "NULL",
                  Predictor != "Intercept", Predictor != "geomorphons",
                  p.value <= 0.05) |> 
    mutate(Predictor = case_when(str_detect(Predictor, "lower_depth") ~ 
                                     str_replace(Predictor,"lower_depth", 
                                                 "Depth"),
                                 str_detect(Predictor, "LITH") ~ 
                                     str_replace(Predictor, "LITH", 
                                                 "PM "),
                                 str_detect(Predictor, "geomorphons") ~
                                     str_replace(Predictor, "geomorphons",
                                                 "Landform "),
                                 str_detect(Predictor, "PET_MAP") ~ 
                                     str_replace(Predictor, "PET_MAP", 
                                                 "PET:MAP "),
                                 str_detect(Predictor, "CHM") ~ 
                                     str_replace(Predictor, "CHM", 
                                                 "Canopy Height"),
                                 str_detect(Predictor, "grad_1000") ~ 
                                     str_replace(Predictor, "grad_1000", 
                                                 "Gradient 1000m"),
                                 str_detect(Predictor, "grad_300") ~ 
                                     str_replace(Predictor, "grad_300", 
                                                 "Gradient 300m"),
                                 str_detect(Predictor, "grad_50") ~ 
                                     str_replace(Predictor, "grad_50", 
                                                 "Gradient 50m"),
                                 str_detect(Predictor, "dev_1000") ~ 
                                     str_replace(Predictor, "dev_1000", 
                                                 "Deviation 1000m"),
                                 str_detect(Predictor, "dev_300") ~ 
                                     str_replace(Predictor, "dev_300", 
                                                 "Deviation 300m"),
                                                .default = (Predictor))) |> 
    dplyr::mutate(Predictor = case_when(str_detect(Predictor, 
                                                   "PM glacial till") ~ 
                                     str_replace(Predictor, 
                                                 "PM glacial till and drift",
                                                 "PM Glacial TD"),
                                     str_detect(Predictor, 
                                                   "PM glacial outwash") ~ 
                                     str_replace(Predictor, 
                                                 "PM glacial outwash",
                                                 "PM Glacial OW"),
                                     str_detect(Predictor, "PM") ~ 
                                     tools::toTitleCase(as.character(Predictor)),
                                     .default = (Predictor)),
                  Predictor = as.factor(Predictor)) |> 
    ggplot(aes(y = fct_reorder(Predictor, -desc(abs(Coefficient))), x = Coefficient)) +
    geom_point(size = 0.7) +
    geom_errorbar(aes(xmin = conf2.5, xmax = conf.975), linewidth = 0.3) + 
    geom_vline(xintercept = 0, lty = 3, colour = "red") +
    scale_x_continuous(name = "Scaled\nCoefficient\nEstimate", n.breaks = 5, 
                       limits = c(-1,1)) +
    scale_y_discrete(labels = scales::label_wrap(15)) +
      theme_param
    #aes(y = stringr::str_wrap(Predictor, 15), x = Estimate) 
    #coord_flip() #+
    
}


```


```{r}
#| fig-width: 6.5
#| fig-height: 6
coef_fig_lmm_stocks <- coef_fig_func(anova_ci_stocks, lmm_stocks) 
coef_fig_lmm_stocks_nw <- coef_fig_func(anova_ci_stocks_nw, lmm_stocks_noWIP) 
coef_fig_lmm_lab <- coef_fig_func(anova_ci_lab, lmm_lab)
coef_fig_lmm_stocks
coef_fig_lmm_stocks_nw
coef_fig_lmm_lab
```

### Random Forest Shap Variable Importance

```{r}
qrf_stocks_shap <- sv_importance(shapviz(ks), kind = "beeswarm", 
                                 max_display = 8, alpha = 0.8) + 
    scale_colour_viridis_c(
      name = "Scaled\nFeature\nValue", option = "D",
      breaks = c(0, 1), labels = c("Low: 0", "High: 1")
    ) +
    scale_y_discrete(labels = \(x) 
    str_replace_all(x, c(lower_depth = "Depth",
                         PET_MAP = "PET:MAP",
                         CHM = "Canopy Height",
                         LITH = "PM",
                         DTM = "Elevation", 
                         geomorphons = "Landform",
                         grad_1000 = "Gradient 1000m",
                         grad_300 = "Gradient 300m",
                         grad_50 = "Gradient 50m",
                         dev_1000 = "Deviation 1000m",
                         dev_300 = "Deviation 300m",
                         dev_50 = "Deviation 50m"))
    ) +
    theme_param 

qrf_stocks_nw_shap <- sv_importance(shapviz(ks_nw), kind = "beeswarm", 
                                 max_display = 8, alpha = 0.8) + 
    scale_colour_viridis_c(
      name = "Scaled\nFeature\nValue", option = "D",
      breaks = c(0, 1), labels = c("Low: 0", "High: 1")
    ) +
    scale_y_discrete(labels = \(x) 
    str_replace_all(x, c(lower_depth = "Depth",
                         PET_MAP = "PET:MAP",
                         CHM = "Canopy Height",
                         LITH = "PM",
                         DTM = "Elevation", 
                         geomorphons = "Landform",
                         grad_1000 = "Gradient 1000m",
                         grad_300 = "Gradient 300m",
                         grad_50 = "Gradient 50m",
                         dev_1000 = "Deviation 1000m",
                         dev_300 = "Deviation 300m",
                         dev_50 = "Deviation 50m"))
    ) +
    theme_param 

qrf_lab_shap <- sv_importance(shapviz(ks_lab), kind = "beeswarm", 
                                 max_display = 8, alpha = 0.8) +
  scale_colour_viridis_c(
      name = "Scaled\nFeature\nValue", option = "D",
      breaks = c(0, 1), labels = c("Low: 0", "High: 1")
    ) +
    scale_y_discrete(labels = \(x) 
        str_replace_all(x, c(lower_depth = "Depth",
                             PET_MAP = "PET:MAP",
                             CHM = "Canopy Height",
                             LITH = "PM",
                             DTM = "Elevation", 
                             geomorphons = "Landform",
                             grad_1000 = "Gradient 1000m",
                             grad_300 = "Gradient 300m",
                             grad_50 = "Gradient 50m",
                             dev_1000 = "Deviation 1000m",
                             dev_300 = "Deviation 300m",
                             dev_50 = "Deviation 50m"))
        ) + theme_param
```

### Full LMM Coefficient and SHAP plots
```{r}
#| fig-width: 6.5
#| fig-height: 6

coef_plots <- (coef_fig_lmm_stocks + ggtitle("Linear Mixed") + ylab(expression(SOC~Stock~g~cm^{2}))) +
  (qrf_stocks_shap) + ggtitle("Random Forest") +
  (coef_fig_lmm_stocks_nw + ylab(expression(SOC~Stock~g~cm^{2})) ) + qrf_stocks_nw_shap +
  (coef_fig_lmm_lab + ylab(expression(SOC~"%"))) +
  qrf_lab_shap + 
    plot_layout(nrow = 3, ncol = 2, axis_titles = "collect", guides = "collect")  +
    plot_annotation(tag_levels = list(c("A.", "B.","C.", "D.", "E.", "F."))) & 
    theme(plot.tag = element_text(size = 8, face = "bold"),
          plot.tag.position = c(0.1, 1.02),
          plot.title = element_text(hjust = 0.5, face = "bold"),
          legend.key.width = unit(0.3, "cm")) 


coef_plots
saveRDS(coef_plots, file = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/LMM_RFM_coef_plots_graphs.rds")

```




```{r}


# save(rfm_pdp_plots, file = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/RFM_PDP_plots.RData")

```


#### Additional Random Forest Shap plots


```{r}
ks_format_func <- function (a_ks){
  ks_vars_form <- a_ks$S |> as.data.frame() |> 
    dplyr::summarise(across(everything(), ~ mean(abs(.x)))) |>
    pivot_longer(everything(), names_to = "column", values_to = "abs_sum") |>
    arrange(desc(abs_sum)) |>
    slice_head(n = 5)
  return(ks_vars_form)
}
ks_vars <- ks_format_func(ks)
ks_wet_vars <- ks_format_func(ks_wet)
ks_upl_vars <- ks_format_func(ks_upl)
ks_nw_vars <- ks_format_func(ks_nw)
ks_full_vars <- ks_format_func(ks_full)

shap_est_func <- function(ks_df, var_name, stockorlab) {
    saveRDS(ks_df[ks_df$column == var_name, "abs_sum"][[1]], 
            paste0(here("writing/in_text_values/"),
                   var_name,
                   "_shap_",
                   stockorlab,
                   ".rds"))
}

for(i in ks_vars$column){
  shap_est_func(ks_vars, i, "stocks")
}


ks_lab_vars <- ks_format_func(ks_lab)
ks_lab_wet_vars <- ks_format_func(ks_lab_wet)
ks_lab_upl_vars <- ks_format_func(ks_lab_upl)

for(i in ks_lab_vars$column){
  shap_est_func(ks_lab_vars, i, "lab")
}


shap_fullmod_plots <- sv_dependence(shapviz(ks), 
                                    v = c(ks_vars$column[1],
                                     ks_vars$column[2], 
                                     ks_vars$column[3], 
                                     ks_vars$column[4]), 
                  color_var = ks_vars$column[5])

shap_fullmod_lab_plots <- sv_dependence(shapviz(ks_lab), 
                                    v = c(ks_lab_vars$column[1],
                                     ks_lab_vars$column[2], 
                                     ks_lab_vars$column[3], 
                                     ks_lab_vars$column[4]), 
                  color_var = ks_lab_vars$column[5])
```


```{r}
ks_wet_vars |> left_join(y = ks_upl_vars, by = join_by(column)) |> 
  dplyr::rename("Wetland_SHAP" = abs_sum.x,
                "Upland_SHAP" = abs_sum.y) |> 
  mutate(WetUplDiff = `Wetland_SHAP` - `Upland_SHAP`) |> 
  mutate(column = fct_relevel(column, c("MAT","MAP", "PET_MAP", "WIP", "lower_depth"))) |>
  ggplot(aes(y = column, x = WetUplDiff)) +
  geom_col() +
    scale_y_discrete(labels = \(x) 
        str_replace_all(x, c(lower_depth = "Depth",
                             PET_MAP = "PET:MAP",
                             CHM = "Canopy Height",
                             LITH = "PM",
                             DTM = "Elevation", 
                             geomorphons = "Landform",
                             grad_1000 = "Gradient 1000m",
                             grad_300 = "Gradient 300m",
                             grad_50 = "Gradient 50m",
                             dev_1000 = "Deviation 1000m",
                             dev_300 = "Deviation 300m",
                             dev_50 = "Deviation 50m"))
        ) + theme_param +
ks_lab_wet_vars |> left_join(y = ks_lab_upl_vars, by = join_by(column)) |> 
  dplyr::rename("Wetland_SHAP" = abs_sum.x,
                "Upland_SHAP" = abs_sum.y) |> 
  mutate(WetUplDiff = `Wetland_SHAP` - `Upland_SHAP`) |> 
  mutate(column = fct_relevel(column, c("PET_MAP","MAP", "Clay", "pH", "lower_depth"))) |>
  ggplot(aes(y = column, x = WetUplDiff)) +
  geom_col() +
    scale_y_discrete(labels = \(x) 
        str_replace_all(x, c(lower_depth = "Depth",
                             PET_MAP = "PET:MAP",
                             CHM = "Canopy Height",
                             LITH = "PM",
                             DTM = "Elevation", 
                             geomorphons = "Landform",
                             grad_1000 = "Gradient 1000m",
                             grad_300 = "Gradient 300m",
                             grad_50 = "Gradient 50m",
                             dev_1000 = "Deviation 1000m",
                             dev_300 = "Deviation 300m",
                             dev_50 = "Deviation 50m"))
        ) + theme_param
```


```{r}
shap_lab_func <- function(shap_plots){
  for(i in 1:length(shap_plots)){
  shap_plots[[i]]$labels$x = str_replace_all(shap_plots[[i]]$labels$x, 
                   c(lower_depth = "Depth",
                     PET_MAP = "PET:MAP",
                     CHM = "Canopy Height",
                     LITH = "PM",
                     DTM = "Elevation",
                     geomorphons = "Landform",
                     grad_1000 = "Gradient 1000m",
                     grad_300 = "Gradient 300m",
                     grad_50 = "Gradient 50m",
                     dev_1000 = "Deviation 1000m",
                     dev_300 = "Deviation 300m",
                     dev_50 = "Deviation 50m"))
    shap_plots[[i]]$labels$colour = str_replace_all(shap_plots[[i]]$labels$colour, 
                   c(lower_depth = "Depth",
                     PET_MAP = "PET:MAP",
                     CHM = "Canopy Height",
                     LITH = "PM",
                     DTM = "Elevation", 
                     geomorphons = "Landform",
                     grad_1000 = "Gradient 1000m",
                     grad_300 = "Gradient 300m",
                     grad_50 = "Gradient 50m",
                     dev_1000 = "Deviation 1000m",
                     dev_300 = "Deviation 300m",
                     dev_50 = "Deviation 50m"))
   shap_plots[[i]]$labels$title = ""
  }
  return(shap_plots)
}
```

Shap partial dependence 

```{r}
shap_1_2_plot <- sv_dependence(shapviz(ks), 
                                    v = c(ks_vars$column[1]), 
                  color_var = ks_vars$column[2])
shap_2_3_plot <- sv_dependence(shapviz(ks), 
                                    v = c(ks_vars$column[3]), 
                  color_var = ks_vars$column[2])
shap_lab_1_2_plot <- sv_dependence(shapviz(ks_lab), 
                                    v = c(ks_lab_vars$column[1]), 
                  color_var = ks_lab_vars$column[2])
shap_lab_2_3_plot <- sv_dependence(shapviz(ks_lab), 
                                    v = c(ks_lab_vars$column[3]), 
                  color_var = ks_lab_vars$column[2])

shap_wipx_depth_plot <- sv_dependence(shapviz(ks), 
                                    v = c(ks_vars$column[2]), 
                  color_var = ks_vars$column[1])

```


```{r}
s <- shapviz(ks_full)
sX <- s$X |> tibble::rowid_to_column(var = "X")

pnw_sX <- left_join(x = sX, y = pnw_dat |> select("X", "site"),  by = join_by("X"))

sd <- sv_dependence(s, v = "WIP", color_var = "lower_depth")
sd_df <- sd$data
pnw_sX$shap <- sd_df$shap

ggplot(data = pnw_sX |> group_by(site), aes(x = WIP, y = shap, colour = site)) +
  geom_line(linewidth = 1)

dep <- tibble(
  feature_val = get_feature_values(s)[["WIP"]]*100,
  shap        = get_shap_values(s)[, "WIP"],
  site        = pnw_dat$site
  ) |> dplyr::mutate(site = str_replace_all(site, c(Hoh = "HRW",
                                                    Mas = "MRW",
                                                    Col = "CVW"))
                     ) 
site_vals <- c("HLEF" = "#a55a1a", "HRW" = "#1d9907", "MRW" = "#007f96", "CVW" = "#f0cf36")
shap_site_plot <- ggplot(dep, aes(feature_val, shap, color = site, fill = site)) +
  geom_point(shape = 21, size = 1.75, colour = "black", alpha = 0.7) +
  geom_line(stat = "smooth", size = 1.1, alpha = 0.8, se = FALSE, method = "loess", formula = y ~ x, show.legend = F) +
  scale_color_manual(values = site_vals) + 
  scale_fill_manual(values = site_vals, name = "Study Area") + 
  labs(x = "WIP %", y = str_c("SHAP Value"), color = "Site") +
  theme_param
shap_site_plot
```


```{r}
shap_pdp_plot_func <- function(shap_interact, name, colour_pal, xlab, reverse = FALSE){
  pal <- RColorBrewer::brewer.pal(9, colour_pal)
  if(reverse) pal <- rev(pal)
  
  x_dat <- shap_interact$data[,2]
  fill_dat <- shap_interact$data[,3]
  if(str_detect(xlab, "WIP")) x_dat <- x_dat*100 
  if(str_detect(name, "WIP")) fill_dat <- fill_dat*100 
  ggplot(shap_interact$data, 
         aes(y = shap, 
             x = x_dat, 
             fill = fill_dat)) +
  geom_point(shape = 21, size = 1.75, colour = "black", alpha = 0.7) + 
  scale_fill_gradientn(name = name, 
                       colours = pal) +
  ylab("SHAP Value") + 
  xlab(xlab) +
  theme_param
}

shap_1_2_pdp_plot <- shap_pdp_plot_func(shap_1_2_plot, 
                                        name = "WIP %", 
                                        colour_pal = "YlGnBu",
                                        xlab = "Depth (cm)")
shap_wipx_depth_pdp_plot <- shap_pdp_plot_func(shap_wipx_depth_plot, 
                                               name = "Depth",
                                               colour_pal = "RdPu",
                                               xlab = "WIP %",
                                               reverse = TRUE)
shap_2_3_pdp_plot <- shap_pdp_plot_func(shap_2_3_plot, 
                                        name = "WIP %", 
                                        colour_pal = "YlGnBu",
                                        reverse = FALSE,
                                        xlab = "PET:MAP")
shap_lab_1_2_pdp_plot <- shap_pdp_plot_func(shap_lab_1_2_plot, 
                                            name = "pH", 
                                        colour_pal = "Reds",
                                        xlab = "Depth (cm)")

shap_lab_2_3_pdp_plot <- shap_pdp_plot_func(shap_lab_2_3_plot, 
                                        name = "pH", 
                                        colour_pal = "Reds",
                                        xlab = "Clay %")
```


```{r}
#| fig-width: 7
#| fig-height: 7
#| out-width: 6.5in
#| fig-dpi: 300
#| 
layout <- "
AAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAA
BBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBB#
CCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCC#
"
shap_site_legend <- ggpubr::get_legend(shap_site_plot)

shap_interact_pdp <- ((shap_site_plot )  + plot_spacer()) /
  (shap_1_2_pdp_plot +  shap_2_3_pdp_plot + plot_layout(guides = "collect", axis_titles = "collect_y") ) /
  (shap_lab_1_2_pdp_plot + shap_lab_2_3_pdp_plot + plot_layout(guides = "collect", axis_titles = "collect_y")) +
  plot_layout(design = layout, nrow = 3, ncol = 1, axis_titles = "collect_y")  + 
  plot_annotation(tag_levels = list(c("A.", "B.","C.", "D.", "E.", "F."))) & 
    theme(plot.tag = element_text(size = 8, face = "bold"),
          plot.tag.position = c(0, 1.02))
shap_interact_pdp 

saveRDS(shap_interact_pdp, file = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/shap_interact_pdp.rds")
```


```{r}
#| message: false
shap_fullmod_plots_form <- shap_lab_func(shap_fullmod_plots)
shap_fullmod_lab_plots_form <- shap_lab_func(shap_fullmod_lab_plots)



shap_full_mod_plot <- shap_fullmod_plots_form + 
    plot_layout(nrow = 2, ncol = 2, axis_titles = "collect", guides = "collect")  +
    plot_annotation(tag_levels = list(c("A.", "B.","C.", "D."))) & 
    theme(plot.tag = element_text(size = 8, face = "bold"),
          plot.tag.position = c(0.1, 1)) &
  theme_param
shap_full_mod_lab_plot <- shap_fullmod_lab_plots_form + 
    plot_layout(nrow = 2, ncol = 2, axis_titles = "collect", guides = "collect")  +
    plot_annotation(tag_levels = list(c("A.", "B.","C.", "D."))) & 
    theme(plot.tag = element_text(size = 8, face = "bold"),
          plot.tag.position = c(0.1, 1)) &
  theme_param

shap_full_mod_plot
shap_full_mod_lab_plot
saveRDS(shap_full_mod_plot, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/shap_full_mod_plot.rds")
saveRDS(shap_full_mod_lab_plot, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/shap_full_mod_lab_plot.rds")

```


```{r}
# ia_1 and ia_lab_1 were saved as already-rendered ggplots (the Interaction object
# was overwritten at the same path upstream) under ggplot2 3.x, and those objects no
# longer dispatch under ggplot2 4.x. Rebuild iml's plot from the stored results
# instead; ia_2-ia_5 are still Interaction objects and keep their own $results.
ia_results <- function(x) {
  as.data.frame(if(is.null(x$results)) x$data else x$results)
}

ia_plot_build <- function(x, sort = TRUE) {
  res <- ia_results(x)
  if(sort) {
    res$.feature <- factor(res$.feature, levels = res$.feature[order(res$.interaction)])
  }
  ggplot(res, aes(y = .feature, x = .interaction)) +
    geom_point() +
    geom_segment(aes(yend = .feature, x = 0, xend = .interaction)) +
    scale_x_continuous("Overall interaction strength") +
    scale_y_discrete("Features")
}

# Shared x axis for a set of panels, taken from the data and rounded up to the next
# 0.05. A hard-coded limit both wastes panel width and silently drops any point
# above it. Each assembled figure sets its own limit from the panels it contains;
# ia_xlim spans all ten and is the fallback for a panel shown on its own.
ia_lim <- function(...) {
  c(0, ceiling(20 * max(sapply(list(...), \(x) max(ia_results(x)$.interaction)))) / 20)
}

ia_xlim <- ia_lim(ia_1, ia_2, ia_3, ia_4, ia_5,
                  ia_lab_1, ia_lab_2, ia_lab_3, ia_lab_4, ia_lab_5)

lab_func <-  function(x, xlim = ia_xlim) {
  subtitle_lab = str_replace_all(
    str_extract(as.character(x$data$.feature), 
                ":lower_depth|:grad_1000|:WIP|:MAT|:Clay|:PET_MAP|:pH|:MAP")[1],
                           c(`:lower_depth` = "Depth",
                             `:grad_1000` = "Gradient 1000m",
                             `:WIP` = "WIP",
                             `:MAT` = "MAT",
                             `:Clay` = "Clay",
                             `:PET_MAP` = "PET::MAP",
                             `:pH` = "pH",
                             `:MAP` = "MAP"
                             ))
  x + scale_y_discrete(labels = \(y) 
        str_replace_all(y, c(`:lower_depth` = "",
                             `:grad_1000` = "",
                             `:WIP` = "",
                             `:MAT` = "",
                             `:Clay` = "",
                             `:PET_MAP` = "",
                             `:pH` = "",
                             `:MAP` = "",
                             lower_depth = "Depth",
                             PET_MAP = "PET::MAP",
                             CHM = "Canopy Height",
                             LITH = "PM",
                             DTM = "Elevation", 
                             geomorphons = "Landform",
                             grad_1000 = "Gradient 1000m",
                             grad_300 = "Gradient 300m",
                             grad_50 = "Gradient 50m",
                             dev_1000 = "Deviation 1000m",
                             dev_300 = "Deviation 300m",
                             dev_50 = "Deviation 50m",
                             SiltClay = "Silt+Clay"
                             )),
        ) + 
    ggtitle(subtitle_lab) +
    scale_x_continuous(limits = xlim) +
    ylab("") +
    xlab("Interaction Strength")
  }

ia_1_plot <- ia_plot_build(ia_1) |> lab_func() + theme_param
ia_2_plot <- plot(ia_2)|> lab_func() + theme_param
ia_3_plot <- plot(ia_3)|> lab_func() + theme_param
ia_4_plot <- plot(ia_4)|> lab_func() + theme_param
ia_5_plot <- plot(ia_5)|> lab_func() + theme_param

ia_lab_1_plot <- ia_plot_build(ia_lab_1) |> lab_func() + theme_param
ia_lab_2_plot <- plot(ia_lab_2)|> lab_func() + theme_param
ia_lab_3_plot <- plot(ia_lab_3)|> lab_func() + theme_param
ia_lab_4_plot <- plot(ia_lab_4)|> lab_func() + theme_param
ia_lab_5_plot <- plot(ia_lab_5)|> lab_func() + theme_param


```

#### Interaction plots from Random Forest 
```{r}
#| fig-width: 6.5
#| out-width: "6.5"
#| message: false

ia_plots <- ia_1_plot + ia_2_plot + ia_3_plot + ia_4_plot +
  plot_layout(nrow = 1, ncol = 4, axis_titles = "collect", guides = "collect")  +
    plot_annotation() & 
    theme(plot.tag = element_text(size = 8, face = "bold"),
          plot.tag.position = c(0.1, 1)) &
  theme_param &
  scale_x_continuous(limits = ia_lim(ia_1, ia_2, ia_3, ia_4))

ia_plots

ia_lab_plots <- ia_lab_1_plot + ia_lab_2_plot + ia_lab_3_plot + ia_lab_4_plot +
  plot_layout(nrow = 1, ncol = 4, axis_titles = "collect", guides = "collect")  +
    plot_annotation() & 
    theme(plot.tag = element_text(size = 8, face = "bold"),
          plot.tag.position = c(0.1, 1)) &
  theme_param &
  scale_x_continuous(limits = ia_lim(ia_lab_1, ia_lab_2, ia_lab_3, ia_lab_4))

ia_lab_plots

ia_both_plots <- (ia_1_plot + ylab(expression(SOC~Stock~g~cm^{2}))) + 
  ia_2_plot + ia_3_plot + 
  (ia_lab_1_plot+ ylab("SOC %")) + ia_lab_2_plot + ia_lab_3_plot +
  plot_layout(nrow = 2, ncol = 3, axis_titles = "collect", guides = "collect")  +
    plot_annotation() & 
    theme(plot.tag = element_text(size = 8, face = "bold"),
          plot.tag.position = c(0.1, 1)) &
  theme_param &
  scale_x_continuous(limits = ia_lim(ia_1, ia_2, ia_3, ia_lab_1, ia_lab_2, ia_lab_3))

ia_both_plots

```


```{r}
saveRDS(ia_plots, file = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/MLR3_RFM_Interact_plot.rds")
saveRDS(ia_lab_plots, file = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/MLR3_RFM_Interact_Lab_plot.rds")
saveRDS(ia_both_plots, file = "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/MLR3_RFM_Interact_BothStocksPerc_plot.rds")

```



LMER Interactions




```{r}
pnw_dat_lab |> filter(WIP < 0.50) |> 
  dplyr::select(carbon_perc, lower_depth, PET_MAP, pH, Clay) |> 
  ppcor::pcor(method = "spearman")

ggplot(data = pnw_dat_lab |> filter(WIP >= 0.50)) +
  geom_point(aes(x = Clay, y = carbon_perc, size = lower_depth, colour = PET_MAP))

```


```{r}
library(interactions)

lmm_stocks_wip_depth_plot <- interact_plot(lmm_stocks,
              pred = lower_depth,
              modx = WIP,
              data = train_dat_scale, jitter = 0,
              plot.points = TRUE,
              colors = "Blues", 
              y.label = expression(log[10] ~ "SOC Stock g cm"^{"-2"}),
              x.label = "Scaled Depth",
              legend.main = "Scaled\nWIP",
              point.alpha = 1)

lmm_lab_ph_petmap_plot <- interact_plot(lmm_lab,
              pred = pH, 
              modx = PET_MAP,
              data = trainlab_scale, jitter = 0,
              plot.points = TRUE,
              colors = "Reds", 
              y.label = expression(log[10] ~ "Mineral SOC %"),
              x.label = "Scaled pH",
              legend.main = "Scaled\nPET:MAP",
              point.alpha = 1)

lmm_interact_plots <- lmm_stocks_wip_depth_plot + lmm_lab_ph_petmap_plot + 
  plot_annotation(tag_levels = "A", tag_suffix = ".") +
  plot_layout() &
  theme(legend.position='top', 
        plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 8, face = "bold"))
lmm_interact_plots
saveRDS(lmm_interact_plots, "/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/SOIL CARBON/All_WA/writing/Figures/lmm_interact_plots.rds")
```

### RF Partial Dependence

Individual Conditional Expectation (ICE)
```{r}
library(mlr3verse)
var <- "WIP"
grid_n <- 50
holdout <- pnw_dat
grid_vals <- holdout |> 
  pull(.data[[var]]) |>
  (\(x) seq(min(x, na.rm = TRUE), max(x, na.rm = TRUE), length.out = grid_n))()

ice <- holdout |>
  mutate(.id = row_number()) |>
  select(-all_of(var)) |>
  cross_join(tibble("{var}" := grid_vals))


ice$pred <- QRFM_stocks$predict_newdata(ice)$response
# average within site at each grid value -> site-level PDP
pdp <- ice |>
  group_by(site, .data[[var]]) |>
  summarise(pred = mean(pred), .groups = "drop")
```


```{r}
ggplot(pdp, aes(.data[[var]], pred, color = site)) +
  geom_line(linewidth = 1) +
  labs(x = var, y = "Predicted response", color = "Site")

ggplot() +
  geom_line(data = ice,
            aes(.data[[var]], pred, group = interaction(site, .id), color = site),
            alpha = 0.1) +
  geom_line(data = pdp,
            aes(.data[[var]], pred, color = site), linewidth = 1.2)

ggplot() +
  geom_violin(data = ice, aes(x = lower_depth, y = pred, color = WIP, group = as.factor(lower_depth)))
```

```{r}

```

### Dead text section

Model choice influences interpretation of drivers and potential upscaling performance

Our choice to use both a LMM and RFM were driven by specifically testing hypothetical model predictors in the LMM while gaining additional insight into potential non-linear predictor relationships using machine learning with RFM. While the LMM, did not perform as well as the RFM on the test datasets, we found the LMM more interpretable regarding potential drivers. This is a well documented issue for machine learning algorithms and it is a challenge to transfer the observations derived from machine learning to process based models or ecosystem processes [@xiaResearchChallengesOpportunities2020]. However, the RFM does not require strict assumptions for the response variable distribution which is the case in the LMM and why we used a $log_{10}$ transformation for both SOC stock and SOC %. Overall we found the RFM consistently out-performed the LMM for both SOC stocks and SOC %. In particular, the RFM was able to maintain higher R2 and low RMSE for both wetland and upland observations in the test dataset. The difference between the LMM and RFM in capturing the SOC % in wetlands test dataset could be the result from the model flexibility of the RFM. The importance of model choice has been emphasized before in Yu et al., (2021) where a random forest outperformed a LMM in estimating SOC concentration in relation to climatic and geochemical predictors, consistent with our research results. Our research shows conservative estimates of SOC stocks are produced when LMMs are used to capture the geographic distribution of depth-dependent SOC stocks. Conversely, the complex relationships between SOC, Depth, and other factors can be captured by models such as the RFM, but with the risk of overfitting to observations and over-predicting SOC stocks in mapped predictions.